\textbf{Analyzing Machupo virus-receptor binding by molecular dynamics simulations}

### 2014-01-05 paper.tex
Updates for Claus's suggestions and added the final figure and data.

We used the final state from each equilibrated system to restart another MD simulation. Our steering protocol is fundamentally similar to \citet{Cuendet2008} with slightly different parameter choices. Perhaps the one significant difference lies in our choosing to not use a thermostat or barostat. However, this should not be a problem since we are not trying to calculate the binding free energy by any physically rigorous approach (the Jarzynski inequality being one example). Following equilibration, the final state of each simulation was used to generate a configuration file fixing the $\alpha$-carbon on residues 1, 58, 73-83, 96, 136, 137, 138, and 161 (again with linear numbering) in the hTfR1. These residues were selected as they are far from the binding interface and sufficiently distributed to prevent any orientational motion of the receptor relative to the viral spike protein. The $\alpha$-carbon of all residues (163-318 in linear numbering) in GP1 received an applied force during the simulation. The NAMD convention does not actually apply a force to all $\alpha$-carbon atoms, but rather uses the selection to compute an initial center of mass. Then, during the steering run, the single center of mass point is pulled with the parameters desribed below. We used the same force field parameters (exclude, cutoff, switching, etc.), the same integrator parameters (time step, rigidbonds on, all molecular being wrapped, etc.), and the same particle mesh ewald parameters as in equilibration. Periodic boundary conditions were incorporated as part of the system (as is the convention in NAMD restart) and PME was again used to approximate full system electrostatics.

@@

We proceeded to analyze this system using two molecular dynamics techniques. First, in SMD by using a known spring constant, and pulling with a constant velocity, we could calculate the applied force during protein--protein dissociation \citep{Cuendet2008,Cuendet2011}. A typical averaged force curve comparison can be seen in Figure~\ref{fig:force_curve}. As seen in Figure~\ref{fig:force_curve}, this distance was relatively consistent among mutants. The supplementary movie visually illustrates the separation distance between peptide domains. The quantities maximum applied force and AUC were derived from the force versus distances curves. Their summary statistics are reported in Table~\ref{tab:summary_statistics}. As we are more interested in the phenotypic impact of interface mutations we avoided many of the more physically rigorous, but technically complicated calculations that are possible with SMD \citep{Is2001A,Is2001B}. Indeed, our set up would not have allowed rigorous calculations as we omitted several tools such as a barostat and a thermostat that would be required for an NPT ensemble.

@@

We proceeded to compare our SMD results to that of the standard dual topology FEP approach to calculate relative free energy differences. The correlation between the energetically rigorous FEP and our statistical approach is high. For all 11 complexes tested, the correlation between max force and FEP was r=0.767 at p=0.00576, and the correlation between AUC and FEP was r=0.72 at p=0.01 (figure). As a result of the strong correlation, we will generally refer to the SMD differences only.

@@

\section{Acknowledgements}
This work was carried out using high-performance computing resources provided by the High Performance Computing Center (HPCC) at Texas Tech University at Lubbock \\ (http://www.hpcc.ttu.edu) and the Texas Advanced Computing Center (TACC) at The University of Texas at Austin (http://www.tacc.utexas.edu). The authors acknowledge Bryan Sutton for opening access to the Hrothgar cluster.

@@

@@

\clearpage
\begin {table}[H]
\caption{\label{tab:summary_statistics}Summary statistics for each mutation tested. $\mu_\text{MAF}$ is the mean and $\sigma_\text{MAF}$ is the standard deviation of maximum applied force over all simulations. $\mu_\text{AUC}$ is the mean and $\sigma_\text{AUC}$ is the standard deviation of AUC over all simulations.}
\begin{center}
  \resizebox{11cm}{!} {
    \begin{tabular}{l c c c c}
    \hline
      Mutation &      $\mu_\text{MAF}$ & $\sigma_\text{MAF}$ & $\mu_\text{AUC}$ & $\sigma_\text{AUC}$ \\ \hline
            WT &                772.00 &              154.13 &         198794.7 &            66765.64 \\
         N348A &                799.33 &              154.36 &         202247.3 &            65175.78 \\
         N348K &                731.36 &              151.01 &         179087.5 &            55799.49 \\
         N348W &                732.13 &              131.21 &         192043.2 &            66916.36 \\
         Y211A &                684.03 &              126.69 &         153517.7 &            40889.12 \\ 
         Y211D &                875.44 &              116.46 &         234909.3 &            50494.11 \\ 
         Y211T &                861.17 &              158.48 &         231078.7 &            97569.29 \\
        vR111A &                754.39 &              103.25 &         185896.8 &            46408.74 \\
   N348A/Y211A &                762.96 &              144.81 &         173059.2 &            59898.10 \\
   N348W/Y211A &                617.21 &              163.33 &         147705.6 &            48548.61 \\
  vR111A/Y211A &                780.71 &              135.65 &         189919.1 &            48868.16 \\
    \hline
    \end{tabular}
  }
\end{center}
\end{table}

>>>>

We used the final state from each equilibrated system to restart another MD simulation. Our steering protocol is fundamentally similar to \citet{Cuendet2008} with slightly different parameter choices. Perhaps the one significant difference lies in our choosing to not use a thermostat or barostat. However, this should not be a problem since we are not trying to calculate the binding free energy by any physically rigorous approach (the Jarzynski inequality being one example). Following equilibration, the final state of each simulation was used to generate a configuration file fixing the $\alpha$-carbon on residues 1, 58, 73-83, 96, 136, 137, 138, and 161 (again with linear numbering) in the hTfR1. These residues were selected as they are far from the binding interface and sufficiently distributed to prevent any orientational motion of the receptor relative to the viral spike protein. The center of mass of the $\alpha$-carbons of all residues (163-318 in linear numbering) in GP1 received an applied force during the simulation. The NAMD convention does not actually apply a force to all $\alpha$-carbon atoms, but rather uses the selection to compute an initial center of mass. Then, during the steering run, the single center of mass point is pulled with the parameters desribed below. We used the same force field parameters (exclude, cutoff, switching, etc.), the same integrator parameters (time step, rigidbonds on, all molecular being wrapped, etc.), and the same particle mesh ewald parameters as in equilibration. Periodic boundary conditions were incorporated as part of the system (as is the convention in NAMD restart) and PME was again used to approximate full system electrostatics.

@@

We proceeded to analyze this system using two molecular dynamics techniques. First, in SMD by using a known force constant, and pulling with a constant velocity, we could calculate the applied force during protein--protein dissociation \citep{Cuendet2008,Cuendet2011}. A typical averaged force curve comparison can be seen in Figure~\ref{fig:force_curve}. As seen in Figure~\ref{fig:force_curve}, this distance was relatively consistent among mutants. The supplementary movie visually illustrates the separation distance between peptide domains. The quantities maximum applied force and AUC were derived from the force versus distances curves. Their summary statistics are reported in Table~\ref{tab:summary_statistics}. As we are more interested in the phenotypic impact of interface mutations we avoided many of the more physically rigorous, but technically complicated calculations that are possible with SMD \citep{Is2001A,Is2001B}.

@@

We proceeded to compare our SMD results to that of the standard dual topology FEP approach to calculate relative free energy differences. The correlation between the energetically rigorous FEP and our statistical approach is high. For all 11 complexes tested, the correlation between max force and FEP was r=-0.767 at p=0.005857 (Figure~\ref{fig:fep}), and the correlation between AUC and FEP was r=--0.729 at p=0.011. As a result of the strong correlation, we will generally refer to the SMD differences only.

@@

\section{Acknowledgements}
This work was carried out using high-performance computing resources provided by the High Performance Computing Center (HPCC) at Texas Tech University at Lubbock \\ (http://www.hpcc.ttu.edu) and the Texas Advanced Computing Center (TACC) at The University of Texas at Austin (http://www.tacc.utexas.edu). The authors acknowledge Bryan Sutton for opening access to the Hrothgar cluster. The authors also acknowledge the reviewers Ilan Samish and Matteo Masetti for their helpful comments in preparing this manuscript.

@@

\clearpage
\begin{figure}[p]
\centerline{\includegraphics[width=6in]{figures/fep_mutants.pdf}}
\caption{\label{fig:fep}\title{Max force versus free energy perturbation} Scatter plot of maximum force in SMD versus the relative free energy difference calculated by FEP for all 10 mutants tested plus the WT complex. The WT complex for FEP was simply set to 0.0. The correlation between the two is r=-0.767 with p=0.005857.}
\end{figure}

@@

\clearpage
\begin {table}[H]
\caption{\label{tab:summary_statistics}Summary statistics for each mutation tested. $\mu_\text{MAF}$ is the mean in piconewtons and $\sigma_\text{MAF}$ is the standard deviation of maximum applied force over all simulations. $\mu_\text{AUC}$ is the mean and $\sigma_\text{AUC}$ is the standard deviation of AUC over all simulations. $\Delta G$ is the free energy difference in kcal/mol calculated via FEP by the dual topology paradigm.}
\begin{center}
  \resizebox{15cm}{!} {
    \begin{tabular}{l c c c c c}
    \hline
      Mutation & $\mu_\text{MAF}$ (pN) & $\sigma_\text{MAF}$ & $\mu_\text{AUC}$ & $\sigma_\text{AUC}$ & $\Delta G$ (kcal/mol)\\ \hline
            WT &                772.00 &              154.13 &         198794.7 &            66765.64 &  0.000\\
         N348A &                799.33 &              154.36 &         202247.3 &            65175.78 & -2.149\\
         N348K &                731.36 &              151.01 &         179087.5 &            55799.49 & +3.184\\
         N348W &                732.13 &              131.21 &         192043.2 &            66916.36 & +3.033\\
         Y211A &                684.03 &              126.69 &         153517.7 &            40889.12 & +2.526\\ 
         Y211D &                875.44 &              116.46 &         234909.3 &            50494.11 & -2.760\\ 
         Y211T &                861.17 &              158.48 &         231078.7 &            97569.29 & +0.876\\
        vR111A &                754.39 &              103.25 &         185896.8 &            46408.74 & +0.467\\
   N348A/Y211A &                762.96 &              144.81 &         173059.2 &            59898.10 & +5.203\\
   N348W/Y211A &                617.21 &              163.33 &         147705.6 &            48548.61 & +8.207\\
  vR111A/Y211A &                780.71 &              135.65 &         189919.1 &            48868.16 & -2.441\\
    \hline
    \end{tabular}
  }
\end{center}
\end{table}
